---
title: "Time_Series"
author: "Debjit_KHaskel(438)"
date: "`r Sys.Date()`"
output: word_document
---

## Description of the data:

Here we are given Monthly totals of international airline passengers, 1949 to 1960. Let X_t denote the monthly total of international airline passengers corresponding to the time point t. Where the origin and the state of time is so chosen that t takes up the values 1,2,3,4,...144(i.e Dec 1948,unit 1) We first obtain time series plot of the data

```{r}
plot(AirPassengers)
```

From the plot we observe that an evolutive pattern is present in the data dominated by an upward trend component. seasonal variation is also present.Cyclical Variation is absent. #Since the intensity of the seasonal fluctuation is increasing with increasing trend , a multiplicative model will be more appropriate hence the model under study is: 

X_t = T_t * S_t *I_t




## THE RATIO TO TREND METHOD

### LINEAR MODEL

#### 1.Yearly averages will be obtained to remove several fluctuations

```{r}
x_year=aggregate(AirPassengers,nfrequency = 1,FUN=mean);x
t_year=1:12
plot(x_year)
```

#### 2.plotting yearly yearly averages against time and fitting an appropiate yearly trend eq

```{r}
plot(x_year)
fit_rt = lm(x_year~t_year)
lines(1949:1960,fitted(fit_rt),type="l",col="red")
```

#### 3.Convert the yearly trend equation to monthly trend equation

```{r}
month_tr=function(t){
  fit_rt$coefficients[1] + fit_rt$coefficients[2]*(t+0.5)/12
}
plot(1:144,AirPassengers,type='l',xlab = "months")
lines(1:144,month_tr(6:149),type='l',col='red')
```

Here we get the monthly trend values from yearly averages

#### 4.Then detrend the original series by taking ratios X_t/T_t

```{r}
detrend_rt = AirPassengers/month_tr(6:149)
plot(detrend_rt,main='Detrended for Linear')
dt_rt = matrix(detrend_rt,ncol=12,byrow=T)
```

#### 5.Obtain adjusted Seasonal indices from the detrended data

```{r}
unadj_s_rt = colMeans(dt_rt)
adj_s_rt = unadj_s_rt /mean(unadj_s_rt)
plot(1949:1960,adj_s_rt,type='l',xlab='Time',main="Adjusted_Seasonal Idices")
```

#### 6.Deseasonalize the detrended data by dividing the detrended series by the corresponding adjusted seasonal


Next we obtain the deseasonalized series by dividing the original data by the corrresponding adjusted seasonal indices
This series should ideally contain the trend component and the irregular fluctuation,
From the plot of deseasonaled data t seems that: 
```{r}
deseasonalized_estimate_rt = detrend_rt/adj_s_rt
plot(1:144,deseasonalized_estimate_rt,main = 'Deseasonalized and Detrended For Linear',type='l',xlab='time_months')
```

### QUADRATIC MODEL

#### 1.Yearly averages will be obtained to remove several fluctuations

```{r}
x_year=aggregate(AirPassengers,nfrequency = 1,FUN=mean);x
t_year=1:12
plot(x_year)
```

#### 2.plotting yearly yearly averages against time and fitting an appropiate yearly trend eq

```{r}
fit_rt_quad = lm(x_year~t_year + I(t_year^2))
plot(x_year)
lines(1949:1960,fitted(fit_rt_quad),type="l",col="red")
```

#### 3.Convert the yearly trend equation to monthly trend equation

```{r}
month_tr_quad=function(t){
  fit_rt_quad$coefficients[1] + fit_rt_quad$coefficients[2]*(t+0.5)/12 + fit_rt_quad$coefficients[3]*((t+0.5)/12)^2
}
plot(1:144,AirPassengers,type='l',xlab = "months",main='Monthly tend')
lines(1:144,month_tr_quad(6:149),type='l',col='red')
```

#### 4.Then detrend the original series by taking ratios X_t/T_t

```{r}
detrend_rt_quad = AirPassengers/month_tr_quad(6:149)
plot(detrend_rt_quad)
dt_rt_quad = matrix(detrend_rt_quad,ncol=12,byrow=T)
```

#### 5.Obtain adjusted Seasonal indices from the detrended data

```{r}
unadj_s_rt_quad = colMeans(dt_rt_quad)
adj_s_rt_quad = unadj_s_rt_quad /mean(unadj_s_rt_quad)
plot(1949:1960,adj_s_rt_quad,type='l',xlab='Time',main="Adjusted_Seasonal Idices")
```

#### 6.Deseasonalize the detrended data by dividing the detrended series by the corresponding adjusted seasonal
Next we obtain the deseasonalized series by dividing the original data by the corrresponding adjusted seasonal indices
This series should ideally contain the trend component and the irregular fluctuation,
From the plot of deseasonaled data t seems that:
```{r}
deseasonalized_estimate_rt_quad = detrend_rt_quad/adj_s_rt_quad
plot(1:144,deseasonalized_estimate_rt_quad,main = 'Deseasonalized and Detrended for Quadratic Model',type='l',xlab='time_months')
```

WE SEE THAT THE QUADRATIC MODEL GIVES MUCH GOOD REPRESENTATION OF IRREGULARITY
