---
title: "Time Series Analysis [JohnsonJohnson Data]"
author: "Arkadyuti Paul [432]"
date: "19/07/2025"
output: word_document
---
```{r}
```


```{r}
```

Data description:
Here we have a Quarterly time series data which contains earnings (dollars) per Johnson & Johnson share from year 1960 to 1980.
```{r}
JohnsonJohnson
```
Objective:
Our objective is to observe the time series data and fit 
Observations: 

```{r}
plot(JohnsonJohnson, xlab = "Year", main = "JohnsonJohnson data")
```


We will take moving average to eliminate the seasonal fluctuation
```{r}
#install.packages("forecast")
library(forecast)
moving_average = ma(JohnsonJohnson,order = 4);moving_average
plot(moving_average, main = "Moving average over Year", ylab = "Moving Average", xlab = "Year")
```
Dtrending the original series
```{r}
dtrended_data = JohnsonJohnson / moving_average ; dtrended_data
plot(dtrended_data, main = "Dtrended data over Year", xlab = "Year", ylab = "Dtrended Data ")
```
Obtaining seasonality index
```{r}
unadjusted_seasonality_index = tapply(na.omit(dtrended_data),cycle(na.omit(dtrended_data)),FUN = mean) ; unadjusted_seasonality_index
adjusted_seasonality_index = unadjusted_seasonality_index / mean(unadjusted_seasonality_index) ; adjusted_seasonality_index
plot(adjusted_seasonality_index, type = 'l', main = "Seasonality index over Quarter", xlab = "Quarter", ylab = "Seasonality Index")
```
Dseasonalize the dtrended data
```{r}
dtrended_data[is.na(dtrended_data)] = 0
dtrended_dseasonalized_data = dtrended_data /  rep(adjusted_seasonality_index, length.out = length(dtrended_data)) ; dtrended_dseasonalized_data
```
Ploting the dtrended dseasonalized data
```{r}
plot(dtrended_dseasonalized_data, main = "Irregularities over year", ylab = "Dtrended dseasonalized data", xlab = "Year")
```






# Step1: We will take yearly average to eliminate the seasonal fluctuation
```{r}
estimated_trend = aggregate(JohnsonJohnson, nfrequency = 1, FUN = mean);estimated_trend
```

#Step2: Plotting yearly average against time and fitting a approximate yearly trend equation
```{r}
plot(estimated_trend, type='l', main = "Trend Estimate VS Time", xlab = "Year", ylab = "Trend Estimate")
time_point = 1:21   
fitted_model_1 = lm(estimated_trend ~ time_point)
summary(fitted_model_1)
year_point = 1960 : 1980
plot(estimated_trend, type = 'l', main = "Fitted yearly trend estimate", xlab = "Year", ylab = "Trend Estimate")
lines(year_point, fitted(fitted_model_1), type = 'l', col = 'red')
```


# Step3: Convert the yearly trend equation to monthly trend equation
```{r}
quarter_point = 6 : 89
quarterly_estimate = function(x){
  fitted_model_1$coefficients[1] + fitted_model_1$coefficients[2] * (x + 1/2) / 4
}
quarterly_estimate(quarter_point)
plot(quarter_point,JohnsonJohnson,type = 'l', main = "Fitted quarterly trend estimate", xlab = "Quarter")
lines(quarter_point,quarterly_estimate(quarter_point),type = 'l', col = 'red')
```


# Step4:Dtrending the original series by taking ratio Xt/Tt
```{r}
dtrended_data_1 = JohnsonJohnson / quarterly_estimate(quarter_point); dtrended_data_1
plot(dtrended_data_1, xlab = "Year", ylab = "Dtrended data", main = "Dtrended data over year")
```


# Step5: Obtaining adjusted seasonal index from the dtrended data 
```{r}
unadjusted_seasonality_index = tapply(dtrended_data_1, cycle(dtrended_data_1), mean) ; unadjusted_seasonality_index
adjusted_seasonality_index = unadjusted_seasonality_index / mean(unadjusted_seasonality_index) ; adjusted_seasonality_index
plot(adjusted_seasonality_index, type = 'l', main = "Seasonality index over Quarter", xlab = "Quarter", ylab = "Seasonality Index")
```


# Step6: Obtaining the dtrended dseasonalized data from dividing dtrended data by adjusted seasonality index
```{r}
dtrended_dseasonalized_data_1 = dtrended_data_1 / rep(adjusted_seasonality_index, length.out = length(dtrended_data_1));dtrended_dseasonalized_data_1
```


# Step7: Plotting this dtrended dseasonalized data 
```{r}
plot(dtrended_dseasonalized_data_1, main = "Irregularities over year", xlab = "Year", ylab = "Dtrended dseasonalized data")
```


###### For quadratic model ######


# Step1: We will take yearly average to eliminate the seasonal fluctuation
```{r}
estimated_trend = aggregate(JohnsonJohnson, nfrequency = 1, FUN = mean);estimated_trend
```

#Step2: Plotting yearly average against time and fitting a approximate yearly trend equation
```{r}
plot(estimated_trend, type='l', main = "Trend Estimate VS Time", xlab = "Year", ylab = "Trend Estimate")
time_point = 1:21   
fitted_model_2 = lm(estimated_trend ~ time_point + I(time_point^2))
summary(fitted_model_2)
year_point = 1960 : 1980
plot(estimated_trend, type = 'l', main = "Fitted yearly trend estimate", xlab = "Year", ylab = "Trend Estimate")
lines(year_point, fitted(fitted_model_2), type = 'l', col = 'red')
```


# Step3: Convert the yearly trend equation to quarterly trend equation
```{r}
quarter_point = 6 : 89
quarterly_estimate = function(x){
  fitted_model_2$coefficients[1] + fitted_model_2$coefficients[2] * (x + 1/2) / 4 + fitted_model_2$coefficients[3] * ((x + 1/2) / 4)^2
}
quarterly_estimate(quarter_point)
plot(quarter_point, JohnsonJohnson, type = 'l', main = "Fitted quarterly trend estimate", xlab = "Quarter")
lines(quarter_point, quarterly_estimate(quarter_point), type = 'l', col = 'red')
```


# Step4:Dtrending the original series by taking ratio Xt/Tt
```{r}
dtrended_data_2 = JohnsonJohnson / quarterly_estimate(quarter_point); dtrended_data_2
plot(dtrended_data_2, xlab = "Year", ylab = "Dtrended data", main = "Dtrended data over year")
```


# Step5: Obtaining adjusted seasonal index from the dtrended data 
```{r}
unadjusted_seasonality_index = tapply(dtrended_data_2, cycle(dtrended_data_2), mean) ; unadjusted_seasonality_index
adjusted_seasonality_index = unadjusted_seasonality_index / mean(unadjusted_seasonality_index) ; adjusted_seasonality_index
plot(adjusted_seasonality_index, type = 'l', main = "Seasonality index over Quarter", xlab = "Quarter", ylab = "Seasonality Index")
```


# Step6: Obtaining the dtrended dseasonalized data from dividing dtrended data by adjusted seasonality index
```{r}
dtrended_dseasonalized_data_2 = dtrended_data_2 / rep(adjusted_seasonality_index, length.out = length(dtrended_data_2));dtrended_dseasonalized_data_2
```


# Step7: Plotting this dtrended dseasonalized data 
```{r}
plot(dtrended_dseasonalized_data_2, main = "Irregularities over year", xlab = "Year", ylab = "Dtrended dseasonalized data")
```

```{r}
CO2
differences=diff(co2,differences = 2)
plot(differences)

differences2=diff(co2,lag= 12)
plot(differences2)

detrend= diff(co2)
deseasonalize= diff(detrend, lag=12)
plot(deseasonalize)

```

```{r}
differences1=diff(AirPassengers)
plot(differences1)

detrend_air= diff(AirPassengers, differences = 2)
deseasonalize_air= diff(detrend_air, lag=12)
plot(deseasonalize_air)

log_detrend_air= diff(log(AirPassengers), differences = 2)
log_deseasonalize_air= diff(log_detrend_air, lag=12)
plot(log_deseasonalize_air)
```
```{r}
plot(log(JohnsonJohnson))
log_detrend_Johnson= diff(log(JohnsonJohnson), differences = 2)
log_deseasonalize_Johnson= diff(log_detrend_Johnson, lag=12)
plot(log_deseasonalize_Johnson)
```
```{r}
#set.seed(1234)
W= rnorm(1000,0,2)
plot(W, type= "l")
Cor= acf(W,lag.max = 50)
Cor
plot(Cor,ci=0.99)
```
```{r}
Q= matrix(c(3,0,1,0,3,1,1,1,2),nrow= 3, byrow= TRUE)
eigen(Q)
d0= c(1,0,0)
Q%*%d0

```
```{r}
WN= rnorm(1000)
plot(WN,type =  "l")

xt= array(0)

for(i in 5:length(WN))
{
  xt[i-4]= (WN[i]+WN[i-1]+WN[i-2]+WN[i-3]+WN[i-4])/5
}

plot(xt,type= "l")

acf(WN)
acf(xt)
xt_filter= filter(WN,filter= rep(1/5,5),sides = 1)
xt_filter= na.omit(xt_filter)
```
```{r}
                                                                                       

sma_ma= ma(white_noise,order = 3)
sma_ma= na.omit(sma_ma)
acf(sma_ma)
```

```{r}
w= rnorm(1000)
xt_ar= filter(w, filter = -0.8, method = "recursive",init = rnorm(1))

acf(xt_ar, lag.max = 50)
pacf(xt_ar, lag.max = 50)

w[1]
xt_ar[1]

xt_ar2= filter(w, filter = c(1,-0.8), method = "recursive",init = rnorm(2))
acf(xt_ar2, lag.max = 50)
pacf(xt_ar2, lag.max = 50)
```
```{r}
set.seed(123)
white_noise= rnorm(500)
delta1=0.5
delta2= 0
delta3= -0.5
xt1=array(0)
xt2=array(0)
xt3=array(0)

xt1[1]=delta1+ white_nose[1]
xt2[1]=delta2+ white_nose[1]
xt3[1]=delta3+ white_nose[1]

for (i in 2:500)
{
  xt1[i]= delta1 + xt1[i-1]+ white_noise[i]
  xt2[i]= delta2 + xt2[i-1]+ white_noise[i]
  xt3[i]= delta3 + xt3[i-1]+ white_noise[i]
}

plot(xt1,type= "l", col= "blue", ylim= c(-250,250))
lines(xt2, type = "l", col= "green")
lines(xt3, type = "l", col= "red")
```
```{r}
acf(co2,lag.max = 100)
pacf(co2, lag.max = 100)

acf(AirPassengers,lag.max = 100)
pacf(AirPassengers, lag.max = 100)
```

```{r}
WN= rnorm(1000)
sma2= filter(WN,filter= rep(1/3,3),sides = 1)
sma2= na.omit(xt_filter)
ar2= filter(WN, filter = c(1,-0.8), method = "recursive",init = rnorm(2))
ar1= filter(WN, filter = 1, method = "recursive",init = rnorm(1))

acf(sma2,lag.max= 50, main= "Simple Moving average of order 2")
acf(ar1,lag.max=349, main= "AR(1) Process")                                                                        
```
```{r}
coefficients <- c(1, 0.9, -0.06)  # Coefficients in increasing order: constant, x, x^2, x^3
roots <- polyroot(coefficients)
print(abs(roots))
```
```{r}
xt_arima= arima.sim(n=5000, list(ar= 0.5, ma=c(1,0.6,0.2)), sd= sqrt(2))
ts.plot(xt_arima)
acf(xt_arima, lag.max= 100)
pacf(xt_arima, lag.max= 100)
```
```{r}
acf(co2,lag.max = 100)
pacf(co2,lag.max = 100)
plot(log(AirPassengers))
Data= log(AirPassengers)
plot(diff(Data))
acf(diff(Data),lag.max = 100)

data=diff(gnp,differences = 1)
plot(data)
acf(data, lag.max = 100)
pacf(data,lag.max = 100)

acf(diff(AirPassengers), lag.max= 100)
```

```{r}
plot(log(gnp))
acf(log(gnp),lag.max = 100)
pacf(log(gnp),lag.max = 100)
data1 = diff(log(gnp),differences = 1)
plot(data1)
acf(data1,lag.max = 100)
pacf(data1,lag.max = 100)
adf.test(data1)
```
```{r}
model1 = arima(log(gnp), order= c(0,1,1));model1
model2= arima(log(gnp), order= c(0,1,2));model2
model3= arima(log(gnp), order= c(1,1,0));model3
model4= arima(log(gnp), order= c(1,1,1));model4


aics= model1$aic
aics= c(aics, model2$aic)
aics= c(aics, model3$aic)
aics= c(aics, model4$aic)
model= c("model 1","model 2","model 3","model 4")
aics
table = data.frame(Models = model, AIC = aics); table

residual1=model1$residuals
residual2=model2$residuals
residual3=model3$residuals
residual4=model4$residuals

par(mfrow= c(2,2))
acf(residual1)
acf(residual2)
acf(residual3)
acf(residual4)

```

```{r}
library(astsa)
data = cardox
index = c(1:floor(0.7*length(data)))
train = data[index]
test = data[-index]
length(train)
length(test)

plot(data)
plot(train,type = 'l')

acf(train,lag.max = 100)
pacf(train,lag.max = 100)

plot(diff(train,differences = 3),type = 'l')

acf(diff(train,differences = 3),lag.max = 100)
pacf(diff(train,differences = 3),lag.max = 100)

a3 = arima(train,c(1,1,0))
a4 = arima(train,c(2,1,0))

acf(a3$residuals,lag.max = 100)
acf(a4$residuals)

#choice for 2nd diff q = 0,1 p = 12 d = 2
#choice for 3rd diff q = 0,1 p = 12 d = 3
model1 = arima(train,c(12,1,1),method = 'ML',optim.control = list(maxit=1500));model1
model2 = arima(train,c(12,2,1),method = 'ML',optim.control = list(maxit=1500));model2
model3 = arima(train,c(12,3,1),method = 'ML',optim.control = list(maxit=1500));model3
model4 = arima(train,c(12,2,2),method = 'ML',optim.control = list(maxit=1500));modle4



par(mfrow = c(1,2))
forecast1 = forecast::forecast(model1,h=length(test))
forecast2= forecast::forecast(model2,h=length(test))
forecast3= forecast::forecast(model3,h=length(test))
forecast4= forecast::forecast(model4,h=length(test))

plot(forecast1)
point_forecast1 = as.numeric(forecast1$mean)
plot(point_forecast1,type = 'l',ylim = c(360,450))
lines(test,col = 'red')


plot(forecast2)
point_forecast2 = as.numeric(forecast2$mean)
plot(point_forecast2,type = 'l',ylim = c(360,450))
lines(test,col = 'red')

plot(forecast3)
point_forecast3 = as.numeric(forecast3$mean)
plot(point_forecast3,type = 'l',ylim = c(360,450))
lines(test,col = 'red')

plot(forecast4)
point_forecast4 = as.numeric(forecast4$mean)
plot(point_forecast4,type = 'l',ylim = c(360,450))
lines(test,col = 'red')

rmse1= sqrt(mean((point_forecast1-test)^2));rmse1
rmse2= sqrt(mean((point_forecast2-test)^2));rmse2
rmse3= sqrt(mean((point_forecast3-test)^2));rmse3
rmse4= sqrt(mean((point_forecast4-test)^2));rmse1
models= c("ARIMA(12,1,1)","ARIMA(12,2,1)","ARIMA(12,3,1)","ARIMA(12,2,2)")
aics= c(model1$aic,model2$aic,model3$aic,model4$aic)
rmses= c(rmse1,rmse2,rmse3,rmse4)
table= data.frame(Models= models, AIC= aics, RMSE= rmses);table

# Box.test(res2,lag=22, type= "Ljung-Box", fitdf = 13)

checkresiduals(model1, lag= 20)
checkresiduals(model2, lag= 20)
checkresiduals(model3, lag= 20)
checkresiduals(model4, lag= 20)



```



```{r}
Data = read.csv("C:\\Users\\DS-31\\Downloads\\data (2).csv")
Sales_ts = ts(Data$Sales, start = c(2000, 1), frequency = 12)

index = 1:floor(0.8 * length(Sales_ts))
train = Sales_ts[index]
test = Sales_ts[-index]

plot(train, type = 'l', main = "Training Sales Data")
acf(train,lag.max = 100)
pacf(train,lag.max = 100)

plot(diff(train,differences = 2),type = 'l')
acf(diff(train,differences = 2),lag.max = 100)
pacf(diff(train,differences = 2),lag.max = 100)


#choice for 2nd diff q = 0,1 p = 0,1,2 d = 2
model1 = arima(train,c(0,2,0),method = 'ML',optim.control = list(maxit=1500))
model2 = arima(train,c(0,2,1),method = 'ML',optim.control = list(maxit=1500))
model3 = arima(train,c(1,2,0),method = 'ML',optim.control = list(maxit=1500))
model4 = arima(train,c(1,2,1),method = 'ML',optim.control = list(maxit=1500))
model5 = arima(train,c(2,2,0),method = 'ML',optim.control = list(maxit=1500))
model6 = arima(train,c(2,2,1),method = 'ML',optim.control = list(maxit=1500))




forecast1 = forecast::forecast(model1,h=length(test))
forecast2= forecast::forecast(model2,h=length(test))
forecast3= forecast::forecast(model3,h=length(test))
forecast4= forecast::forecast(model4,h=length(test))
forecast5= forecast::forecast(model4,h=length(test))
forecast6= forecast::forecast(model4,h=length(test))

par(mfrow = c(1,2))
plot(forecast1)
point_forecast1 = as.numeric(forecast1$mean)
plot(point_forecast1,type = 'l',ylim = c(5000,10000))
lines(test,col = 'red')


plot(forecast2)
point_forecast2 = as.numeric(forecast2$mean)
plot(point_forecast2,type = 'l',ylim = c(5000,10000))
lines(test,col = 'red')

plot(forecast3)
point_forecast3 = as.numeric(forecast3$mean)
plot(point_forecast3,type = 'l',ylim = c(5000,10000))
lines(test,col = 'red')

plot(forecast4)
point_forecast4 = as.numeric(forecast4$mean)
plot(point_forecast4,type = 'l',ylim = c(5000,10000))
lines(test,col = 'red')

plot(forecast5)
point_forecast5 = as.numeric(forecast5$mean)
plot(point_forecast5,type = 'l',ylim = c(5000,10000))
lines(test,col = 'red')

plot(forecast6)
point_forecast6 = as.numeric(forecast6$mean)
plot(point_forecast6,type = 'l',ylim = c(5000,10000))
lines(test,col = 'red')

rmse1= sqrt(mean((point_forecast1-test)^2))
rmse2= sqrt(mean((point_forecast2-test)^2))
rmse3= sqrt(mean((point_forecast3-test)^2))
rmse4= sqrt(mean((point_forecast4-test)^2))
rmse5= sqrt(mean((point_forecast5-test)^2))
rmse6= sqrt(mean((point_forecast6-test)^2))

models= c("ARIMA(0,2,0)","ARIMA(0,2,1)","ARIMA(1,2,0)","ARIMA(1,2,1)","ARIMA(2,2,0)","ARIMA(2,2,1)")
aics= c(model1$aic,model2$aic,model3$aic,model4$aic,model5$aic,model6$aic)
rmses= c(rmse1,rmse2,rmse3,rmse4,rmse5,rmse6)
table= data.frame(Models= models, AIC= aics, RMSE= rmses);table


checkresiduals(model1, lag= 20)
checkresiduals(model2, lag= 20)
checkresiduals(model3, lag= 20)
checkresiduals(model4, lag= 20)
checkresiduals(model5, lag= 20)
checkresiduals(model6, lag= 20)
```

```{r}
auto.arima(train)
```

```{r}
GNP= log(gnp)
index = 1:floor(0.8 * length(GNP))
train = GNP[index]
test = GNP[-index]

plot(train, type = 'l', main = "Training Sales Data")
acf(train,lag.max = 100)
pacf(train,lag.max = 100)

plot(diff(train,differences = 1),type = 'l')
acf(diff(train,differences = 1),lag.max = 100)
pacf(diff(train,differences = 1),lag.max = 100)


#choice for 2nd diff  p = 0,1 q = 0,1,2 d = 1
model1 = arima(train,c(0,1,0),method = 'ML',optim.control = list(maxit=1500))
model2 = arima(train,c(0,1,1),method = 'ML',optim.control = list(maxit=1500))
model3 = arima(train,c(0,1,2),method = 'ML',optim.control = list(maxit=1500))
model4 = arima(train,c(1,1,0),method = 'ML',optim.control = list(maxit=1500))
model5 = arima(train,c(1,1,1),method = 'ML',optim.control = list(maxit=1500))
model6 = arima(train,c(1,1,2),method = 'ML',optim.control = list(maxit=1500))




forecast1 = forecast::forecast(model1,h=length(test))
forecast2= forecast::forecast(model2,h=length(test))
forecast3= forecast::forecast(model3,h=length(test))
forecast4= forecast::forecast(model4,h=length(test))
forecast5= forecast::forecast(model4,h=length(test))
forecast6= forecast::forecast(model4,h=length(test))

par(mfrow = c(1,2))
plot(forecast1)
point_forecast1 = as.numeric(forecast1$mean)
plot(point_forecast1,type = 'l',ylim = c(5000,10000))
lines(test,col = 'red')


plot(forecast2)
point_forecast2 = as.numeric(forecast2$mean)
plot(point_forecast2,type = 'l',ylim = c(5000,10000))
lines(test,col = 'red')

plot(forecast3)
point_forecast3 = as.numeric(forecast3$mean)
plot(point_forecast3,type = 'l',ylim = c(5000,10000))
lines(test,col = 'red')

plot(forecast4)
point_forecast4 = as.numeric(forecast4$mean)
plot(point_forecast4,type = 'l',ylim = c(5000,10000))
lines(test,col = 'red')

plot(forecast5)
point_forecast5 = as.numeric(forecast5$mean)
plot(point_forecast5,type = 'l',ylim = c(5000,10000))
lines(test,col = 'red')

plot(forecast6)
point_forecast6 = as.numeric(forecast6$mean)
plot(point_forecast6,type = 'l',ylim = c(5000,10000))
lines(test,col = 'red')

rmse1= sqrt(mean((point_forecast1-test)^2))
rmse2= sqrt(mean((point_forecast2-test)^2))
rmse3= sqrt(mean((point_forecast3-test)^2))
rmse4= sqrt(mean((point_forecast4-test)^2))
rmse5= sqrt(mean((point_forecast5-test)^2))
rmse6= sqrt(mean((point_forecast6-test)^2))

models= c("ARIMA(0,1,0)","ARIMA(0,1,1)","ARIMA(0,1,2)","ARIMA(1,1,0))","ARIMA(1,1,1)","ARIMA(1,1,2)")
aics= c(model1$aic,model2$aic,model3$aic,model4$aic,model5$aic,model6$aic)
rmses= c(rmse1,rmse2,rmse3,rmse4,rmse5,rmse6)
table= data.frame(Models= models, AIC= aics, RMSE= rmses);table


checkresiduals(model1, lag= 20)
checkresiduals(model2, lag= 20)
checkresiduals(model3, lag= 20)
checkresiduals(model4, lag= 20)
checkresiduals(model5, lag= 20)
checkresiduals(model6, lag= 20)
```

